[Re] Non-additive coupling enables propagation of synchronous spiking activity in purely random networks

### 2018-03-13 article/Caze_2018.tex
update the ve wording

The originality of this model lays in a filtering of the excitatory inputs by the function $\sigma$ to model dendritic non-linearities. The study considers two functions $\sigma$ (see Fig.~1 of the original paper and insets in Fig.~\ref{fig:markov_corrected}). Linear coupling, defined by $\sigma(x)=x$, and non-linear coupling with:
\begin{equation}
    \sigma(x)= 
\begin{cases}
    x              & \text{, if } x\leq V_a\\
    V_a + \frac{V_c-V_a}{V_b-V_a} (x - V_a), & \text{, if } V_a<x \leq V_b\\
    V_c            & \text{, otherwise.}
\end{cases}
\end{equation}
For this non-linear function, the biological interpretation is that below a first threshold $V_a$, excitatory inputs are propagated passively by the dendrite and sum linearly. Above this threshold, they trigger a dendritic spike, resulting in an amplified somatic voltage. If they are even higher and cross the threshold $V_b$, they stop increasing the somatic voltage and saturate. Note that this non-linearity only affects spikes received synchronously as has been observed experimentally. With the introduction of this non-linearity, the modelled point neuron becomes an effective model of a neuron with one non-linear dendrite.

@@

For non-linearly coupled networks, we have to use a slightly more complicated approach where we sum up all the received excitatory input in a variable \verb|ve|, apply the non-linear function $\sigma$ to this input, and then reset the variable \verb|ve| back to 0 to prepare for the next time step. In Brian syntax (the definition of the inhibitory synapses is identical to the linearly coupled network):
\begin{code}
eq = """dV/dt = -gamma*V + V0 : volt
        ve : volt
     """
G = br2.NeuronGroup(N, model=eq, threshold="V>Theta",
                    reset="V=0*mV", method='euler')
exc_syn = br2.Synapses(G, G, 'w : volt (constant, shared)',
                       on_pre='ve += w', delay=5 * ms - dt)
G.run_regularly('''
                V += clip(ve, 0*mV, 2*mV) + clip(2*(ve-2*mV), 0*mV, 4*mV)
                ve = 0*mV
                ''', when='after_synapses')
\end{code}
The \verb|run_regularly| operation that is run every time step to implement the non-linear $\sigma$ function refers to the predefined \verb|clip| function which clips the value to a given range, following the syntax
\begin{code}
clip(value, lower_bound, upper_bound)
\end{code}

@@

{\sffamily \small
   \printbibliography[title=References]
%}
\clearpage
\appendix
\section*{Supplementary Figures}
\begin{figure}[h]
\centering
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_l_dt_100us_gs100.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs100.pdf}
\end{subfigure}
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs75.pdf}
\end{subfigure}
\caption{\label{fig:grid_max}
\textbf{Qualitative network behaviour as a function of excitatory and inhibitory connection strength for linearly (A) and non-linearly (B, C) coupled networks.} All conventions as in Fig.~\ref{fig:grid}, but points are coloured according the observed qualitative behaviour that occurred the most often among the 10 independent simulations for given connection strengths.}
\end{figure}

\begin{figure}[h]
\centering
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs100_mean_detail.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_50us_gs100_mean_detail.pdf}
\end{subfigure}
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_10us_gs100_mean_detail.pdf}
\end{subfigure}
\caption{\label{fig:grid_detail}
\textbf{Qualitative network behaviour as a function of excitatory and inhibitory connection strength for non-linearly coupled networks simulated with different simulation time steps.} Color code as in Fig.~\ref{fig:grid}. Results are shown for a smaller range of connection strengths than in Fig.~\ref{fig:grid}, and only for non-linearly coupled networks with an synchronous stimulation of 100 neurons. The total inhibitory connection strength was varied between \SI{24}{\milli\volt} and \SI{34.55}{\milli\volt}, the total excitatory connection strength was varied between \SI{34.91}{\milli\volt} and \SI{45.45}{\milli\volt}, i.e. the parameter range corresponds to a square region at the top of Fig.~\ref{fig:grid}B. Simulations where performed with simulation time steps of \SI{0.1}{\milli\second} (A; also the value used for all other simulations shown), \SI{0.05}{\milli\second} (B), and \SI{0.01}{\milli\second} (C).}
\end{figure}

\begin{figure}
\centering
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/spikes_nl_strong_dt_100us.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/spikes_nl_strong_dt_10us.pdf}
\end{subfigure}
\caption{\label{fig:raster_strong}
\textbf{Propagation of synchronous spiking in non-linearly coupled networks with strong recurrent connections when simulated with a time step of \SI{0.1}{\milli\second} (A), or \SI{0.01}{\milli\second} (B).} All conventions as in Fig.~\ref{fig:raster}, but for strong recurrent connections ($w_\text{ex} = w_\text{in} = \SI{0.4}{\milli\volt}$). Both simulated networks use non-linearly coupling, the network shown in panel (B) is simulated with a 10-times finer temporal resolution compared to the other results shown.}
\end{figure}

\begin{figure}[h]
\centering
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_l_dt_100us.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_nl_dt_100us.pdf}
\end{subfigure}
\caption{\label{fig:markov}
\textbf{Evolution of synchronous pulses in linearly (A) and non-linearly (B) coupled networks, corrected for the background activity.} All conventions as in Fig.~\ref{fig:markov_corrected}. The numerical results shown here (gray lines and green squares) are not corrected for the bias introduced by the clock-driven integration method (see text for details).}
\end{figure}
\end{document}

>>>>

The originality of this model lays in a filtering of the excitatory inputs by the function $\sigma$ to model dendritic non-linearities. The study considers two functions $\sigma$ (see Fig.~1 of the original paper and insets in Fig.~\ref{fig:markov_corrected}). Linear coupling, defined by $\sigma(x)=x$, and non-linear coupling with:
\begin{equation}
    \sigma(x)= 
\begin{cases}
    x              & \text{, if } x\leq V_a\\
    V_a + \frac{V_c-V_a}{V_b-V_a} (x - V_a), & \text{, if } V_a<x \leq V_b\\
    V_c            & \text{, otherwise.}
\end{cases}
\label{eq:non-linear}
\end{equation}
For this non-linear function, the biological interpretation is that below a first threshold $V_a$, excitatory inputs are propagated passively by the dendrite and sum linearly. Above this threshold, they trigger a dendritic spike, resulting in an amplified somatic voltage. If they are even higher and cross the threshold $V_b$, they stop increasing the somatic voltage and saturate. Note that this non-linearity only affects spikes received synchronously as has been observed experimentally. With the introduction of this non-linearity, the modelled point neuron becomes an effective model of a neuron with one non-linear dendrite.

@@

We used a the predefined \verb|clip| function to implemented non-linear coupling into a network. The function clips a value to a given range, following the syntax
\begin{code}
clip(value, lower_bound, upper_bound)
\end{code}

We employed this functions on the received excitatory inputs stored in a placed holder variable \verb|ve|. The variable corresponds to x in eq.~\ref{eq:non-linear}. We then reset the variable \verb|ve| back to 0 to prepare for the next time step. In Brian syntax (the definition of the inhibitory synapses is identical to the linearly coupled network):

\begin{code}
eq = """dV/dt = -gamma*V + V0 : volt
        ve : volt
     """
G = br2.NeuronGroup(N, model=eq, threshold="V>Theta",
                    reset="V=0*mV", method='euler')
exc_syn = br2.Synapses(G, G, 'w : volt (constant, shared)',
                       on_pre='ve += w', delay=5 * ms - dt)
G.run_regularly('''
                V += clip(ve, 0*mV, 2*mV) + clip(2*(ve-2*mV), 0*mV, 4*mV)
                ve = 0*mV
                ''', when='after_synapses')
\end{code}

The \verb|run_regularly| operation enables that to run the non-linear summation between every time steps.

@@

{\sffamily \small
   \printbibliography[title=References]
%}
\clearpage
\appendix
\section*{Supplementary Figures}
\begin{figure}[h]
\centering
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_l_dt_100us_gs100.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs100.pdf}
\end{subfigure}
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs75.pdf}
\end{subfigure}
\caption{\label{fig:grid_max}
\textbf{Using max instead of mean to color each pixels.} All conventions as in Fig.~\ref{fig:grid}, but points are coloured according the observed qualitative behaviour that occurred the most often among the 10 independent simulations for given connection strengths.}
\end{figure}

\begin{figure}[h]
\centering
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_100us_gs100_mean_detail.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_50us_gs100_mean_detail.pdf}
\end{subfigure}
\begin{subfigure}[t]{0.33\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/grid_nl_dt_10us_gs100_mean_detail.pdf}
\end{subfigure}
\caption{\label{fig:grid_detail}
\textbf{Zoom on qualitative network behaviour with different simulation time steps.} Color code as in Fig.~\ref{fig:grid}. Results are shown for a smaller range of connection strengths than in Fig.~\ref{fig:grid}, and only for non-linearly coupled networks with an synchronous stimulation of 100 neurons. The total inhibitory connection strength was varied between \SI{24}{\milli\volt} and \SI{34.55}{\milli\volt}, the total excitatory connection strength was varied between \SI{34.91}{\milli\volt} and \SI{45.45}{\milli\volt}, i.e. the parameter range corresponds to a square region at the top of Fig.~\ref{fig:grid}B. Simulations where performed with simulation time steps of \SI{0.1}{\milli\second} (A; also the value used for all other simulations shown), \SI{0.05}{\milli\second} (B), and \SI{0.01}{\milli\second} (C).}
\end{figure}

\begin{figure}
\centering
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/spikes_nl_strong_dt_100us.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/spikes_nl_strong_dt_10us.pdf}
\end{subfigure}
\caption{\label{fig:raster_strong}
\textbf{Propagation of synchronous spiking when simulated with small timesteps}.(A) Time step of \SI{0.1}{\milli\second}, or \SI{0.01}{\milli\second} (B) All conventions as in Fig.~\ref{fig:raster}, but for strong recurrent connections ($w_\text{ex} = w_\text{in} = \SI{0.4}{\milli\volt}$). Both simulated networks use non-linearly coupling, the network shown in panel (B) is simulated with a 10-times finer temporal resolution compared to the other results shown.}
\end{figure}

\begin{figure}[h]
\centering
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_l_dt_100us.pdf}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\caption{}\vspace{-1ex}
\includegraphics[width=\textwidth]{Figs/markov_nl_dt_100us.pdf}
\end{subfigure}
\caption{\label{fig:markov}
\textbf{Evolution of synchronous pulses when using smaller timesteps in linearly (A) and non-linearly (B) coupled networks, corrected for the background activity.} All conventions as in Fig.~\ref{fig:markov_corrected}. The numerical results shown here (gray lines and green squares) are not corrected for the bias introduced by the clock-driven integration method (see text for details).}
\end{figure}
\end{document}
